---
title: "Hotel booking prediction"
format: 
  html: 
    toc: true
editor: visual
execute: 
  warning: false 
  output: false 
---

This document takes the code found in this step-by-step instruction for the *tidymodels* package and formats it as Quarto document for practice purposes.

## Loading packages

First, attach the packages required for the following approaches.

```{r packageLoading}
library(tidymodels)  
library(readr)       # for importing data
library(vip)         # for variable importance plots
```

## Loading and analyzing data

Loading of the hotel bookings dataset.

```{r dataLoading}

hotels <- 
  read_csv("https://tidymodels.org/start/case-study/hotels.csv", show_col_types = FALSE) %>%
  mutate(across(where(is.character), as.factor))
```

Depicting the dataset dimension below.

```{r dataDimensions}
#| output: true
dim(hotels)
```

Of interest to predict are the stays with children. Let's check how many there are in the complete dataset.

```{r numberOfChildren}
#| output: true
hotels %>% 
  count(children) %>% 
  mutate(prop = n/sum(n))
```

8.1% is a rather low number and the dataset hence misbalanced. Nevertheless, we will proceed with the data as-is.

## Data splitting and resampling

Let's split the data into training and test data.

```{r}
set.seed(123)
splits      <- initial_split(hotels, strata = children)

hotel_other <- training(splits)
hotel_test  <- testing(splits)
```

These are the training test proportions of stays with children:

```{r}
#| output: true
hotel_other %>% 
  count(children) %>% 
  mutate(prop = n/sum(n))
```

These are the test test proportions of stays with children:

```{r}
#| output: true 
hotel_test %>%  
  count(children) %>%    
  mutate(prop = n/sum(n))
```

We can see that the test proportions of stays with children are accurately reflected.

The validation set will now be generated:

```{r}
set.seed(234)
val_set <- validation_split(hotel_other, 
                            strata = children, 
                            prop = 0.80)
```

With the data prepared, it it time to apply the models.

## 1st model: penalized logistic regression

Building the model:

```{r}
lr_mod <- 
  logistic_reg(penalty = tune(), mixture = 1) %>% 
  set_engine("glmnet")
```

Create the recipe:

```{r}
holidays <- c("AllSouls", "AshWednesday", "ChristmasEve", "Easter", "ChristmasDay", "GoodFriday", "NewYearsDay", "PalmSunday")

lr_recipe <- 
  recipe(children ~ ., data = hotel_other) %>% 
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, holidays = holidays) %>% 
  step_rm(arrival_date) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
```

Create the workflow:

```{r}
lr_workflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(lr_recipe)
```

Create the tuning grid:

```{r}
#| output: true
lr_reg_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))

lr_reg_grid %>% top_n(-5) # lowest penalty values
lr_reg_grid %>% top_n(5)  # highest penalty values
```

Train and tune the model:

```{r}
lr_res <- 
  lr_workflow %>% 
  tune_grid(val_set,
            grid = lr_reg_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))

```

```{r}
#| output: true
lr_plot <- 
  lr_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number())

lr_plot
```

```{r}
#| output: true
top_models <-
  lr_res %>% 
  show_best(metric = "roc_auc", n = 15) %>% 
  arrange(penalty) 
top_models
```

```{r}
#| output: true
lr_best <- 
  lr_res %>% 
  collect_metrics() %>% 
  arrange(penalty) %>% 
  slice(12)
lr_best
```

```{r}
#| output: true
lr_auc <- 
  lr_res %>% 
  collect_predictions(parameters = lr_best) %>% 
  roc_curve(children, .pred_children) %>% 
  mutate(model = "Logistic Regression")

autoplot(lr_auc)
```

## 2nd model: penalized logistic regression

Building the model:
